---
title: "Injuries in the NFL"
author: "Christopher Spartz"
date: "11/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r setup-libs}
library(rvest)
library(stringr)
library(tidyverse)
library(ggplot2)
library(purrr)
#library(lubridate)
library(xml2)

teamids <- c('mia','nwe','buf','nyj','pit','rav','cin','cle','jax','oti',
           'htx','clt','kan','sdg','rai','den','phi','dal','was','nyg',
           'min','det','gnb','chi','nor','car','atl','tam','ram','sea',
           'crd','sfo')
years <- c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)
```

```{r inj-func}
team_inj <- function(t) {
  c <- 1
  tinj <- list()
  for (y in seq_along(years)) {
    url <- sprintf("https://www.pro-football-reference.com/teams/%s/%d_injuries.htm",t,years[y])
    i <- url %>%
      read_html()%>%
      html_node('table#team_injuries')%>%
      html_table()
    i <- i[c(1:17)]
    names(i) <- c("Player","Game_1","Game_2","Game_3","Game_4","Game_5","Game_6","Game_7","Game_8","Game_9",
              "Game_10","Game_11","Game_12","Game_13","Game_14","Game_15","Game_16")
    i <- i %>% pivot_longer(!c("Player"),names_to = "Game",values_to = "Injury")
    i <- i %>% separate(Game,c(NA,"Game"),"_")
    i <- i[!(is.na(i$Injury) | i$Injury==""), ]
    i <- i %>% mutate(year=years[y],
                  team=t)
    i$Player <- as.factor(i$Player)
    tinj[[c]] <- i
    c <-  c+1
  }
  big_inj <- tinj %>% bind_rows()
}
```
```{r all-inj}
teaminj <- list()
c <- 1
for (t in seq_along(teamids)) {
  teaminj[[c]] <- team_inj(teamids[[t]])
  c <- c+1
}
all_inj <- teaminj %>% bind_rows()
```

```{r}
all_inj <- all_inj %>% filter(Injury!="//")
all_inj <- all_inj %>% filter(Injury!="//;90")
all_inj <- all_inj %>% filter(Injury!="S")
all_inj$Injury <- as.factor(all_inj$Injury)
all_inj$Injury <- factor(all_inj$Injury,levels = c("C19","P","Q","D","O","IR","PUP"))
all_inj %>% group_by(Injury) %>% count(Injury)
```


```{r starter-func}
team_starters <- function(t) {
  url <- sprintf("https://www.pro-football-reference.com/teams/%s/lineups.htm#starting_lineups",t)
  lin <- read_html(url)
  p <- lin %>%
    html_node('table#starting_lineups')
  
  xml_find_all(p, ".//br") %>% xml_add_sibling("p", "\n")

  xml_find_all(p, ".//br") %>% xml_remove()
  p <- p %>% html_table()
  p <- p %>% pivot_longer(!c("Year","Record"),names_to = "Position Group",values_to = "Players") %>% filter(Year>2009)
  p1 <- p %>% separate(Players,c("Player_1","Player_2","Player_3","Player_4","Player_5","Player_6"),sep = "\n")
  p2 <- p1 %>% pivot_longer(!c("Year","Record","Position Group"),names_to = "Field") %>% na.omit()
  p3 <- p2 %>% mutate(value = str_replace(value, "\\s", "|")) %>% 
    separate(value, into = c("Position", "Player"), sep = "\\|")
  p3$"Player" <- str_remove_all(p3$'Player','[*+]')
  p3 <- p3 %>% select(-Field)
  p3 <- p3 %>% mutate(team=t)
}
```

```{r}
teamStart <- list()
c <- 1
for (t in seq_along(teamids)) {
  teamStart[[c]] <- team_starters(teamids[[t]])
  c <- c+1
}
all_starters <- teamStart %>% bind_rows()
all_starters$Player <- as.factor(all_starters$Player)
all_starters$Position <- factor(all_starters$Position,levels = c("QB","RB","FB","WR","TE","LT","LG","C",
                                                                 "RG","RT","LDT","RDT","DE","DT","NT","LDE","RDE",
                                                                 "LOLB","LILB","RILB","ROLB","LLB","MLB","RLB",
                                                                 "LCB","RCB","CB","DB","SS","FS","PR","KR"))
all_starters <- all_starters %>% separate(Record,c("Wins","Losses","Ties"), sep = "-")
all_starters$Wins <- as.numeric(all_starters$Wins)
all_starters$Losses <- as.numeric(all_starters$Losses)
all_starters$Ties <- as.numeric(all_starters$Ties)
```


```{r join-dbs}
all_starters$Player <- str_trim(all_starters$Player)
alltogether <- all_starters %>% rename(year=Year) %>% inner_join(all_inj, by=c("Player","year","team"))
alltogether$team <- as.factor(alltogether$team)
```

Injuries are split into 8 categories these represent:<br />
C19 stands for a player was reported as having had covid 19, this only shows up in last years data, but I thought it was interesting so I kept it in. <br />
P stands for probable, and when a team reports a player as probable that means that he has a 75% chance of playing <br />
Q stands for questionable, and when a team reports a player as questionable that means that he has a 50% chance of playing <br />
D stands for doubtful, and when a team reports a player as doubtful that means that he has a 25% chance of playing <br />
O stands for out, which means that player can't play that week but is not put onto the injured reserve or PUP <br />
IR stands for Injured Reserve, which is a list that teams can put players onto that means that player is unable to play that week and it opens up a spot for an uninjured player to play that week. <br />
PUP means the player is on the physically unable to perform list which means that they cannot play or practice with the team for the first 6 weeks but they can still attend meetings and use the medical/training facilities. <br />


```{r inj-by-year}
library(ggrepel)
all_inj %>% group_by(Injury,year) %>% count(Injury) %>% unique() %>% filter(Injury!="C19") %>%
  mutate(label = if_else(year == 2020, as.character(Injury), NA_character_)) %>%
  ggplot(aes(x=year,y=n, group=Injury , color=Injury)) +
  geom_line() + 
  geom_label_repel(aes(label = label),
                  nudge_x = 1,
                  na.rm = TRUE) +
  ylab("Games with a player placed in each injury Category") +
  ggtitle("Games with a player in an injury group from 2010-2020")
```



```{r inj-by-position, fig.height=10, fig.width=10}
impPos <- alltogether[!(alltogether$Position=="DE" | alltogether$Position=="DT" | 
                          alltogether$Position=="CB" | alltogether$Position=="DB"),]
impPos %>% group_by(Injury,Position) %>% count(Injury) %>%
  ggplot(aes(x=Injury,y=n, fill=Injury)) +
  geom_bar(stat = "Identity") + 
  ggtitle("Injuries by position") +
  ylab("Weeks spent with injuries") +
  facet_wrap(~Position,ncol = 4)
```
In this graph all of the positions before LDE are offensive position, while all of the positions after LDE and including LDE are defensive positions.
This graph shows how some positions, like WR, which are the positions that 


```{r winpercent}
alltogether %>% group_by(team,year) %>% 
  mutate(teaminjs=n(), winp=(Wins+0.5*Ties)/16) %>%
  ggplot(aes(x=winp,y=teaminjs, color=team)) + 
  geom_point() +
  geom_smooth(method=lm , color="black", se=FALSE) +
  facet_wrap(~year)
```


This graph shows 

```{r inj-win-team, fig.height=6, fig.width=10}
alltogether %>% group_by(team,year) %>% 
  mutate(teaminjs=n(), winp=(Wins+0.5*Ties)/16) %>%
  ggplot(aes(x=winp,y=teaminjs,color=team))  +
  geom_smooth(method=lm , color="black", se=FALSE) +
  geom_point() + facet_wrap(~team,ncol = 8)
```







